---
title: "Activity Data Analysis"
author: "Fan Zheng"
date: "June 6, 2016"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r load data, echo = TRUE, cache=TRUE}
library(ggplot2)
stepdata<-read.csv("activity.csv", head=TRUE,na.strings="NA")
```

## What is mean total number of steps taken per day?

```{r , echo=TRUE, cache=TRUE}
totalperday<-tapply(stepdata$steps,stepdata$date,sum,na.rm=TRUE)
hist(totalperday,col="green")
```

## Mean and Median of the total number of steps taken per day
Mean:
```{r mean of total number of steps taken per day, echo = TRUE, cache=TRUE}
meanperday<-mean(totalperday)
meanperday
```
Median:
```{r median of total number of steps taken per day, echo=TRUE, cache=TRUE}
medianperday<-median(totalperday)
medianperday
```
## What is the average daily activity pattern?
Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days 
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r time series plot of steps vs interval, echo=TRUE, cache=TRUE}
intervalperday<-tapply(stepdata$steps,stepdata$interval,mean,na.rm=TRUE)
plot(rownames(intervalperday),intervalperday)
```

Max interval:
```{r max interval and index, echo=TRUE, cache=TRUE}
maxinterval<-which.max(intervalperday)
names(intervalperday[maxinterval])
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset.
2. Devise a strategy for filling in all of the missing values in the dataset. Here I use the mean for that 5-minute          interval over all days.
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
   Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number    of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the      impact of imputing missing data on the estimates of the total daily number of steps?

Number of NAs:

```{r number of NAs, echo=TRUE, cache=TRUE}
sum(is.na(stepdata$steps))
```

Fill the NAs:
```{r fill the NAs,echo=TRUE, cache=TRUE}
a<-which(is.na(stepdata$steps))
b<-stepdata$interval[which(is.na(stepdata$steps))]
b<-as.character(b)
c<-intervalperday[b]
stepdata1<-stepdata
stepdata1$steps[a]<-c
```

Mean of the new data set without NAs:
```{r echo=TRUE,cache=TRUE}
totalperday1<-tapply(stepdata1$steps,stepdata1$date,sum,na.rm=TRUE)
meanperday1<-mean(totalperday1)
meanperday1
```

Median of the new data set without NAs:
```{r echo=TRUE,cache=TRUE}
medianperday1<-median(totalperday1)
medianperday1
```

Histogram of the new dataset mean total steps per day without NAs (right, orange) compared with the original one (left, green):
```{r echo=TRUE,cache=TRUE}
par(mfrow=c(1,2))
hist(totalperday,col="green")
hist(totalperday1,col="orange")
```

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date     is a weekday or weekend day.
2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps         taken, averaged across all weekday days or weekend days (y-axis).
```{r Create factor variable for weekend and weekdays, echo=TRUE,cache=TRUE}
stepdata1$date<-as.Date(stepdata1$date)
stepdata1$weekdays<-weekdays(stepdata1$date)
weekday <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
weekend <- c('Saturday', 'Sunday')
stepdata1$level<-stepdata1$weekdays
stepdata1[(stepdata1$weekdays) %in% weekday,]$level<-"Weekday"
stepdata1[(stepdata1$weekdays) %in% weekend,]$level<-"Weekend"
```

Panel Plot:
```{r panel plot,ech=TRUE,cache=TRUE}
stepdata1_intervalperday<-aggregate(steps~level+interval,data=stepdata1,mean)
g<-ggplot(stepdata1_intervalperday,aes(interval,steps))
g<-g+geom_line()+facet_wrap(~level,ncol=1)+theme(strip.text.x=element_text(size=10,angle=0),strip.background=element_rect(fill="wheat2"))+
  xlab("Interval")+ylab("Number of steps")
print(g)
```
